---
title: "01_demodata"
output: html_document
---

This R code gathers demographic information for census block groups
in California. It outputs data files containing this aggregated
information.

Census data was downloaded through the Census Bureau (U.S. Department of 
Commerce) website: https://data.census.gov/. The data considered are from 
the "American Community Survey, ACS 5-Year Estimates".

CalEnviroScreen (CES) data used is from CalEnviroScreen 4.0 results.
More information in CES report: https://oehha.ca.gov/media/downloads/calenviroscreen/report/calenviroscreen40reportf2021.pdf


```{r}
library(tidyverse)
library(readxl)
library(stringr)
```

------------------------------------------------------------------------

CENSUS BLOCK GROUPS: DEMOGRAPHIC DATA IMPORT

------------------------------------------------------------------------

The following tables are downloaded via the American FactFinder Download
Center: - B01003 - TOTAL POPULATION - B02001 - RACE - B03002 - HISPANIC
OR LATINO ORIGIN BY RACE - B03003 - HISPANIC OR LATINO ORIGIN - B15003 -
EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER - B19001 -
HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED
DOLLARS) - B19013 - MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN
2021 INFLATION-ADJUSTED DOLLARS) - B25003 - TENURE - B25024 - UNITS IN
STRUCTURE - B25034 - YEAR STRUCTURE BUILT - B25035 - MEDIAN YEAR
STRUCTURE BUILT

```{r}
# B01003 - TOTAL POPULATION
# loading data, renaming columns, removing extra header row
ACS_2021_bg_totpop = read.csv("blockgroup/ACS_21_5YR_B01003_with_ann.csv", header=TRUE, sep = ";", skip=1, na.strings="null",
                              colClasses=c("factor", "factor", "character", "character", "character"))
# Rename the columns to match the expected names
colnames(ACS_2021_bg_totpop) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totpop","totpop_err")
# Remove double quotes and convert to numeric
ACS_2021_bg_totpop <- ACS_2021_bg_totpop %>%
  mutate_at(vars(totpop, totpop_err), ~as.numeric(gsub('"', '', .)))
```

```{r}
# B02001 - RACE, including total population
ACS_2021_bg_race = read.csv("blockgroup/ACS_21_5YR_B02001_with_ann.csv", ,header=TRUE, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character", rep("character",20)))
colnames(ACS_2021_bg_race) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totpop", "totpop_err", "white", "white_err", "black", "black_err", "native", "native_err", "asian", "asian_err", "pacific", "pacific_err", "other", "other_err", "two", "two_err", "twoincl", "twoincl_err", "twoexcl", "twoexcl_err")
# Convert the last 20 columns to numeric and remove double quotes
ACS_2021_bg_race <- ACS_2021_bg_race %>%
  mutate_at(vars(tail(names(.), 20)), ~as.numeric(gsub('"', '', .)))
# creating new summary columns
ACS_2021_bg_race <- ACS_2021_bg_race %>%
  mutate(twoormore = two, #+ twoincl + twoexcl,
         white_pct = white / totpop,
         black_pct = black / totpop, 
         native_pct = native / totpop,
         asian_pct = asian / totpop,
         pacific_pct = pacific / totpop,
         other_pct = other / totpop,
         twoormore_pct = twoormore / totpop) %>%
  mutate(racediversity = 1 - (white*(white-1) + black*(black-1) + native*(native-1) + asian*(asian-1) + pacific*(pacific-1) + 
                                other*(other-1) + twoormore*(twoormore-1))/(totpop*(totpop-1)))
```

```{r}
# B03002 - HISPANIC OR LATINO ORIGIN BY RACE
ACS_2021_bg_latinxrace = read.csv("blockgroup/ACS_21_5YR_B03002_with_ann.csv", header=TRUE, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character", rep("character",42)))
colnames(ACS_2021_bg_latinxrace) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totpop", "totpop_err", "nlx", "nlx_err", "nlxwhite", "nlxw_err", "nlxbblack", "nlxb_err", "nlxnative", "nlxn_err", "nlxasian", "nlxa_err", "nlxpacific", "nlxp_err", "nlxother", "nlxo_err", "nlxtwo", "nlxtwo_err", "nlxtwoincl", "nlxtwoincl_err", "nlxtwoexcl", "nlxtwoexcl_err", "lx", "lx_err", "lxwhite", "lxw_err", "lxbblack", "lxb_err", "lxnative", "lxn_err", "lxasian", "lxa_err", "lxpacific", "lxp_err", "lxother", "lxo_err", "lxtwo", "lxtwo_err", "lxtwoincl", "lxtwoincl_err", "lxtwoexcl", "lxtwoexcl_err")
# Convert the last 42 columns to numeric and remove double quotes
ACS_2021_bg_latinxrace <- ACS_2021_bg_latinxrace %>%
  mutate_at(vars(tail(names(.), 42)), ~as.numeric(gsub('"', '', .)))
# creating new summary columns
ACS_2021_bg_latinxrace <- ACS_2021_bg_latinxrace %>%
  mutate(nlxtwoormore = nlxtwo + nlxtwoincl + nlxtwoexcl,
         lxtwoormore = lxtwo + lxtwoincl + lxtwoexcl,
         nlxwhite_pct = nlxwhite / totpop)
```

```{r}
# B03003 - HISPANIC OR LATINO ORIGIN
ACS_2021_bg_latinx = read.csv("blockgroup/ACS_21_5YR_B03003_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character",rep("character",6)))
colnames(ACS_2021_bg_latinx) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totpop", "totpop_err", "notlatinx", "notlatinx_err", "latinx", "latinx_err")
# Convert the last 6 columns to numeric and remove double quotes
ACS_2021_bg_latinx <- ACS_2021_bg_latinx %>%
  mutate_at(vars(tail(names(.), 6)), ~as.numeric(gsub('"', '', .)))
# creating new summary columns
ACS_2021_bg_latinx <- ACS_2021_bg_latinx %>%
  mutate(notlatinx_pct = notlatinx / totpop,
         latinx_pct = latinx / totpop)
```

```{r}
# B15003 - EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER
# includes total population, but with different (lower) numbers than tot pop or race
ACS_2021_bg_educ = read.csv("blockgroup/ACS_21_5YR_B15003_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character",rep("character",50)))
colnames(ACS_2021_bg_educ) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totpop", "totpop_err", "none", "none_err", "nursery", "nursery_err", "kindergarten", "kindergarten_err", "grd01", "grd01_err", "grd02", "grd02_err", "grd03", "grd03_err", "grd04", "grd04_err", "grd05", "grd05_err", "grd06", "grd06_err", "grd07", "grd07_err", "grd08", "grd08_err", "grd09", "grd09_err", "grd10", "grd10_err", "grd11", "grd11_err", "grd12nodip", "grd12nodip_err", "hsdip", "hsdip_err", "GED", "GED_err", "somecol1", "somecol1_err", "somecol1plsnodeg", "somecol1plsnodeg_err", "associate", "associate_err", "bachelor", "bachelor_err", "master", "master_err", "prof", "prof_err", "phd", "phd_err")
# Convert the last 50 columns to numeric and remove double quotes
ACS_2021_bg_educ <- ACS_2021_bg_educ %>%
  mutate_at(vars(tail(names(.), 50)), ~as.numeric(gsub('"', '', .)))
# creating some new summary columns and percentages
ACS_2021_bg_educ <- ACS_2021_bg_educ %>%
  mutate(presch = none + nursery + kindergarten,
         elem = grd01 + grd02 + grd03 + grd04 + grd05,
         mids = grd06 + grd07 + grd08,
         somehs = grd09 + grd10 + grd11 + grd12nodip,
         hsfin = hsdip + GED,
         somecol = somecol1 + somecol1plsnodeg,
         grad = master + prof + phd) %>% 
  mutate(presch_pct = presch / totpop,
         elem_pct = elem / totpop,
         mids_pct = mids / totpop,
         somehs_pct = somehs / totpop,
         hsnofin_pct = (presch + elem + mids + somehs) / totpop,
         hsfin_pct = hsfin / totpop,
         somecol_pct = somecol / totpop,
         associate_pct = associate / totpop,
         bachelor_pct = bachelor / totpop,
         grad_pct = grad / totpop) %>%
  mutate(edavgyrs = (presch*0 + grd01*1 + grd02*2 + grd03*3 + grd04*4 + grd05*5 + grd06*6 + grd07*7 + grd08*8 + 
                 grd09*9 + grd10*10 + grd11*11 + grd12nodip*11 + hsfin*12 + somecol*14 + associate*14 + bachelor*16 + 
                 master*18 + prof*20 + phd*22)/totpop)
```

```{r}
# B19001 - HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED DOLLARS), including number of households
ACS_2021_bg_hhinc = read.csv("blockgroup/ACS_21_5YR_B19001_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character", rep("numeric",34)))
colnames(ACS_2021_bg_hhinc) <- c("GEOid_bg", "GEOID10", "bgcountystate", "tothh", "tothh_err", "lessthan10k", "lessthan10k_err", "inc10to15k", "inc10to15k_err", "inc15to20k", "inc15to20k_err", "inc20to25k", "inc20to25k_err", "inc25to30k", "inc25to30k_err", "inc30to35k", "inc30to35k_err", "inc35to40k", "inc35to40k_err", "inc40to45k", "inc40to45k_err", "inc45to50k", "inc45to50k_err", "inc50to60k", "inc50to60k_err", "inc60to75k", "inc60to75k_err", "inc75to100k", "inc75to100k_err", "inc100to125k", "inc100to125k_err", "inc125to150k", "inc125to150k_err", "inc150to200k", "inc150to200k_err", "inc200kplus", "inc200kplus_err")
# Convert the last 34 columns to numeric and remove double quotes
ACS_2021_bg_hhinc <- ACS_2021_bg_hhinc %>%
  mutate_at(vars(tail(names(.), 34)), ~as.numeric(gsub('"', '', .)))
# new column for census tract ID, since this data is the starting point for the full dataset
ACS_2021_bg_hhinc$GEOIDCT <- str_sub(ACS_2021_bg_hhinc$GEOID10, 1, -2)
# creating some new summary columns
ACS_2021_bg_hhinc <- ACS_2021_bg_hhinc %>%
  mutate(inc25kbelow = lessthan10k + inc10to15k + inc15to20k + inc20to25k,
         inc25to50k = inc25to30k + inc30to35k + inc35to40k + inc40to45k + inc45to50k,
         inc50to75k = inc50to60k + inc60to75k,
         inc100to150k = inc100to125k + inc125to150k,
         inc50kbelow = inc25kbelow + inc25to50k,
         inc50to100k = inc50to75k + inc75to100k,
         inc100to200k = inc100to150k + inc150to200k) %>%
  mutate(inc25kbelow_pct = inc25kbelow / tothh,
         inc25to50k_pct = inc25to50k / tothh,
         inc50to75k_pct = inc50to75k / tothh,
         inc75to100k_pct = inc75to100k / tothh,
         inc100to150k_pct = inc100to150k / tothh,
         inc150to200k_pct = inc150to200k / tothh,
         inc50kbelow_pct = inc50kbelow / tothh,
         inc50to100k_pct = inc50to100k / tothh,
         inc100to200k_pct = inc100to200k / tothh,
         inc200kplus_pct = inc200kplus / tothh,
         inc150kplus_pct = (inc150to200k + inc200kplus) / tothh)
```

```{r}
# B19013 - MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED DOLLARS)
ACS_2021_bg_medhhinc = read.csv("blockgroup/ACS_21_5YR_B19013_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null")
colnames(ACS_2021_bg_medhhinc) <- c("GEOid_bg", "GEOID10", "bgcountystate", "medhhinc", "medhhinc_err")
ACS_2021_bg_medhhinc$GEOID10 <- str_pad(ACS_2021_bg_medhhinc$GEOID10, width=12, side="left", pad="0")
ACS_2021_bg_medhhinc$GEOID10 <- as.factor(ACS_2021_bg_medhhinc$GEOID10)
ACS_2021_bg_medhhinc$bgcountystate <- as.character(ACS_2021_bg_medhhinc$bgcountystate)
ACS_2021_bg_medhhinc$medhhinc <- gsub("-","",ACS_2021_bg_medhhinc$medhhinc)
ACS_2021_bg_medhhinc$medhhinc <- gsub("[^0-9-]", "", ACS_2021_bg_medhhinc$medhhinc)
ACS_2021_bg_medhhinc$medhhinc <- as.numeric(ACS_2021_bg_medhhinc$medhhinc)
ACS_2021_bg_medhhinc$GEOID10 <- substring(ACS_2021_bg_medhhinc$GEOID10, 2)
```

```{r}
# B25003 - TENURE, including number of households
ACS_2021_bg_tenure = read.csv("blockgroup/ACS_21_5YR_B25003_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character",rep("character",6)))
colnames(ACS_2021_bg_tenure) <- c("GEOid_bg", "GEOID10", "bgcountystate", "tothh", "tothh_err", "ownerocc", "ownerocc_err", "renterocc", "renterocc_err")
# Convert the last 6 columns to numeric and remove double quotes
ACS_2021_bg_tenure <- ACS_2021_bg_tenure %>%
  mutate_at(vars(tail(names(.), 6)), ~as.numeric(gsub('"', '', .)))
# percentages
ACS_2021_bg_tenure <- ACS_2021_bg_tenure %>%
  mutate(ownerocc_pct = ownerocc / tothh,
         renterocc_pct = renterocc / tothh)
```

```{r}
# B25024 - UNITS IN STRUCTURE, including number of structures
ACS_2021_bg_units = read.csv("blockgroup/ACS_21_5YR_B25024_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor", "factor", "character",rep("character",22)))
colnames(ACS_2021_bg_units) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totstr", "totstr_err", "units1det", "units1det_err", "units1att", "units1att_err", "units2", "units2_err", "units3to4", "units3to4_err", "units5to9", "units5to9_err", "units10to19", "units10to19_err", "units20to49", "units20to49_err", "units50plus", "units50plus_err", "mobilehome", "mobilehome_err", "boatrvvan", "boatrvvan_err")
# Convert the last 22 columns to numeric and remove double quotes
ACS_2021_bg_units <- ACS_2021_bg_units %>%
  mutate_at(vars(tail(names(.), 22)), ~as.numeric(gsub('"', '', .)))
# creating some new summary columns
ACS_2021_bg_units <- ACS_2021_bg_units %>%
  mutate(singleunit = units1att + units1det,
         duplex = units2,
         smallmulti = units3to4 + units5to9 + units10to19,
         largemulti = units20to49 + units50plus) %>%
  mutate(singleunit_pct = singleunit / totstr,
         duplex_pct = duplex / totstr,
         smallmulti_pct = smallmulti / totstr,
         largemulti_pct = largemulti / totstr) %>%
  mutate(MUD_pct = duplex_pct + smallmulti_pct + largemulti_pct) %>%
  mutate(unitsavg = (singleunit*1 + duplex*2 + units3to4*3.5 + units5to9*7 + units10to19*14.5 + units20to49*34.5 + units50plus*50) / totstr)
```

```{r}
# B25034 - YEAR STRUCTURE BUILT, including number of structures
ACS_2021_bg_yrbuilt = read.csv("blockgroup/ACS_21_5YR_B25034_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null", colClasses=c("factor","factor","character",rep("character",22)))
colnames(ACS_2021_bg_yrbuilt) <- c("GEOid_bg", "GEOID10", "bgcountystate", "totstr", "totstr_err", "yr2014plus", "yr2014plus_err", "yr2010to2013", "yr2010to2013_err", "yr2000to2009", "yr2000to2009_err", "yr1990to1999", "yr1990to1999_err", "yr1980to1989", "yr1980to1989_err", "yr1970to1979", "yr1970to1979_err", "yr1960to1969", "yr1960to1969_err", "yr1950to1959", "yr1950to1959_err", "yr1940to1949", "yr1940to1949_err", "yr1939pre", "yr1939pre_err")
# Convert the last 22 columns to numeric and remove double quotes
ACS_2021_bg_yrbuilt <- ACS_2021_bg_yrbuilt %>%
  mutate_at(vars(tail(names(.), 22)), ~as.numeric(gsub('"', '', .)))
# creating some new summary columns
ACS_2021_bg_yrbuilt <- ACS_2021_bg_yrbuilt %>%
  mutate(yr2000post = yr2014plus + yr2010to2013 + yr2000to2009,
         yr1980to1999 = yr1990to1999 + yr1980to1989,
         yr1960to1979 = yr1970to1979 + yr1960to1969,
         yr1959pre = yr1950to1959 + yr1940to1949 + yr1939pre) %>%
  mutate(yr2000post_pct = yr2000post / totstr,
         yr1980to1999_pct = yr1980to1999 / totstr,
         yr1960to1979_pct = yr1960to1979 / totstr,
         yr1959pre_pct = yr1959pre / totstr)
```

```{r}
# B25035 - MEDIAN YEAR STRUCTURE BUILT
ACS_2021_bg_medyrbuilt = read.csv("blockgroup/ACS_21_5YR_B25035_with_ann.csv", header=T, sep = ";", skip=1, na.strings="null")
colnames(ACS_2021_bg_medyrbuilt) <- c("GEOid_bg", "GEOID10", "bgcountystate", "medyrbuilt", "medyrbuilt_err")
ACS_2021_bg_medyrbuilt$medyrbuilt <- gsub("[^0-9-]", "", ACS_2021_bg_medyrbuilt$medyrbuilt)
ACS_2021_bg_medyrbuilt <- ACS_2021_bg_medyrbuilt %>%
  mutate(GEOID10 = as.factor(str_pad(GEOID10, width=12, side="left", pad="0")),
         bgcountystate = as.character(bgcountystate),
         medyrbuilt = as.numeric(gsub("-","",medyrbuilt)))
ACS_2021_bg_medyrbuilt$GEOID10 <- substring(ACS_2021_bg_medyrbuilt$GEOID10, 2)

```

------------------------------------------------------------------------

CENSUS BLOCK GROUPS: AGGREGATING CENSUS DATA

------------------------------------------------------------------------

```{r}
geomatch = c("GEOid_bg", "GEOID10", "bgcountystate")


bg_demo <- ACS_2021_bg_hhinc[ , c(geomatch, "GEOIDCT", "tothh", "inc50kbelow_pct", "inc150kplus_pct")]

bg_demo <- bg_demo %>% # HOUSEHOLD-LEVEL METRICS, tothh from income data
  merge(ACS_2021_bg_medhhinc[, c(geomatch, "medhhinc")], by=geomatch, all.x=TRUE) %>% # median household income
  merge(ACS_2021_bg_tenure[, c(geomatch, "ownerocc_pct", "renterocc_pct")], by=geomatch, all.x=TRUE) %>% # tenure
  merge(ACS_2021_bg_race[, c(geomatch, "totpop", "racediversity", "black_pct", "asian_pct")],
        by=geomatch, all.x=TRUE) %>% # POPULATION-LEVEL METRICS, totpop from race data
  merge(ACS_2021_bg_latinxrace[, c(geomatch, "nlxwhite_pct")], by=geomatch, all.x=TRUE) %>% # non-hispanic white
  merge(ACS_2021_bg_latinx[, c(geomatch, "notlatinx_pct", "latinx_pct")], by=geomatch, all.x=TRUE) %>% # latinx
  merge(ACS_2021_bg_educ[, c(geomatch, "edavgyrs")], by=geomatch, all.x=TRUE) %>% # education
  merge(ACS_2021_bg_units[, c(geomatch, "totstr", "singleunit_pct", "MUD_pct", "duplex_pct", "smallmulti_pct", "largemulti_pct",  "unitsavg")], 
        by=geomatch, all.x=TRUE) %>% # STRUCTURE-LEVEL METRICS, totstr from unit data
  merge(ACS_2021_bg_yrbuilt[, c(geomatch, "yr2000post_pct", "yr1980to1999_pct", "yr1960to1979_pct", "yr1959pre_pct")], 
        by=geomatch, all.x=TRUE) %>% # years built
  merge(ACS_2021_bg_medyrbuilt[, c(geomatch, "medyrbuilt")], by=geomatch, all.x=TRUE) # median year built
```

```{r}
rm(list=ls(pattern="ACS_2021_bg"))
```

------------------------------------------------------------------------

CENSUS TRACTS: CALENVIROSCREEN

------------------------------------------------------------------------

```{r}
# read in CalEnviroScreen 4.0 results from xlsx file
# while importing, convert Census Tract GEOID to characters to match with block group ID
CES4results = read_excel("CalEnviroScreen/ces4results.xlsx", sheet=1, na="NA", col_types=c("text","numeric",rep("guess",56)))
#colnames(CES4results)[colnames(CES4results)=="Census Tract"] <- "GEOIDCT"
colnames(CES4results) <- c("GEOIDCT", "totpop", "county", "zip_ct", "nearbycity", "longitude", "latitude", "ces4_sc", "ces4_pctl", "ces4_pctlrng", "ozone", "ozone_pctl", "pm25", "pm25_pctl", "dieselpm", "dieselpm_pctl", "drinkingwtr", "drinkingwtr_pctl", "lead", "lead_pctl", "pesticides", "pesticides_pctl", "toxrelease", "toxrelease_pctl", "traffic", "traffic_pctl", "cleanupsites", "cleanupsites_pctl", "grndwtrthreats", "grndwtrthreats_pctl", "hazwaste", "hazwaste_pctl", "impwtrbodies", "impwtrbodies_pctl", "solidwaste", "solidwaste_pctl", "polburden", "polburden_sc", "polburden_pctl", "asthma", "asthma_pctl", "lowbirthwt", "lowbirthwt_pctl", "cardiovdis", "cardiovdis_pctl", "educ", "educ_pctl", "lingisolation", "lingisolation_pctl", "poverty", "poverty_pctl", "unemployment", "unemployment_pctl", "housingburden", "housingburden_pctl", "popchar", "popchar_sc", "popchar_pctl")

# prepend 0 to every census tract GEOID to enable matching with block group IDs
#CES4results$GEOIDCT <- str_pad(CES4results$GEOIDCT, width=11, side="left", pad="0")

CES4results <- CES4results %>%
  mutate(polexposure_pctl = rowMeans(select(., c("ozone_pctl", "pm25_pctl", "dieselpm_pctl", "pesticides_pctl", 
                                                 "toxrelease_pctl", "traffic_pctl", "drinkingwtr_pctl")), na.rm=TRUE),
         polenvt_pctl = rowMeans(select(., c("cleanupsites_pctl", "grndwtrthreats_pctl", "hazwaste_pctl", 
                                             "impwtrbodies_pctl", "solidwaste_pctl")), na.rm=TRUE),
         popsens_pctl = rowMeans(select(., c("asthma_pctl", "lowbirthwt_pctl", "cardiovdis_pctl")), na.rm=TRUE),
         popsocioecon_pctl = rowMeans(select(., c("educ_pctl", "lingisolation_pctl", "poverty_pctl", "unemployment_pctl", "housingburden_pctl")), na.rm=TRUE))
```

------------------------------------------------------------------------

CENSUS BLOCK GROUPS: AGGREGATING CENSUS AND CALENVIROSCREEN DATA

------------------------------------------------------------------------

```{r}
# Tract vs block group matching, https://www.census.gov/geo/reference/geoidentifiers.html
# 11 characters for census tract, 12 (1 addtl) for block group
# merge CES4 results for percentile (?) with demographic data spreadsheet
bg_demo <- bg_demo %>%
  merge(CES4results[ , c("GEOIDCT", "county", "zip_ct", "nearbycity", "polexposure_pctl", "polenvt_pctl", "popsens_pctl", "popsocioecon_pctl", "lingisolation_pctl", "polburden_pctl", "popchar_pctl", "ces4_sc", "ces4_pctl")],
        by=c("GEOIDCT"), all.x=TRUE)
```

```{r}
demotypes <- sapply(bg_demo,class)
write.csv(demotypes, file = "bgCAcensusCESclasses.csv", row.names=FALSE)
write.csv(bg_demo, file="bgCAcensusCES.csv", row.names=FALSE)
```
