---
title: "05_access"
output: html_notebook
---

This notebook calculates grid access for households in PG&E and SCE service territories

VERSION N°3 - General calculations for household access

```{r}
library(tidyverse)
library(tictoc)
library(reshape2)
library(ggplot2)
library(stringr)
library(tidytext)
library(measurements) # https://cran.r-project.org/web/packages/measurements/measurements.pdf
library(readxl)
library(colorspace)
library(cowplot)
```

```{r}
setwd('C:/Users/Louis Rosset/Documents/EPFL/Master Project/QGIS_data/04_allocation/IOUdata/SCE')
SCEtypes = read.csv("circuitfiles/SCE_ICAalldemoallocclasses.csv", header=T, na.strings="?")
SCE_ICAaccess = read.csv("circuitfiles/SCE_ICAalldemoalloc.csv", header=T, na.strings="?", colClasses=as.character(SCEtypes$x))

setwd('C:/Users/Louis Rosset/Documents/EPFL/Master Project/QGIS_data/04_allocation/IOUdata/PGE')
PGEtypes = read.csv("circuitfiles/PGE_ICAalldemoallocclasses.csv", header=T, na.strings="?")
PGE_ICAaccess = read.csv("circuitfiles/PGE_ICAalldemoalloc.csv", header=T, na.strings="?", colClasses=as.character(PGEtypes$x))

#Correcting some duplicates problem
PGE_ICAaccess <- PGE_ICAaccess[!duplicated(PGE_ICAaccess[, c("GEOID", "CircuitName")]), ]

at <- c("1", "1.5", "2.0", "3.0", "4.0", "4.5", "5.0", "6.0", "7.0", "8.0", "9.0", "10.")
```

```{r}
labsIOU <- function(x) ifelse(x == "PGE", "PG&E", "SCE")
labsDER <- function(x) ifelse(x == "ICL", "Load", NA)
```

--------------------------------------------------------------------------------------------

GRID LIMITS

--------------------------------------------------------------------------------------------

```{r}
plotops_bggrid = list(
  theme_light(), facet_grid(IOU~DER, scales="free_x", space='free_x', labeller=labeller(IOU=labsIOU, DER=labsDER)),
  geom_boxplot(aes(fill=threshold), alpha=1, width=0.9), scale_fill_discrete_sequential(palette="Mint", nmax=13, order=1:13), 
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), legend.position="bottom"), guides(fill=guide_legend(nrow=1)), 
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1), labels=scales::percent_format(accuracy=1)), 
  labs(x="", y="Portion of households with access, by block group", fill="Access threshold (kW)")
)

plotops_allgrid = list(theme_light(), facet_grid(IOU~DER, labeller=labeller(IOU=labsIOU)),
  geom_line(), geom_point(size=2), 
  theme(legend.position="bottom", legend.margin=margin(), legend.text=element_text(size=8), axis.text.x=element_text(size=7)),
  guides(shape=guide_legend(nrow=4,title.position="top")),
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1), labels=scales::percent_format(accuracy=1)),
  scale_x_continuous(breaks=c(0,as.numeric(at)), labels=c(">0", "1", "1.5","2","3","4","4.5","5","6","7","8","9","10")),
  scale_color_discrete_sequential(palette="Viridis", nmax=5, order=5:2),
  labs(x="Access threshold (kW)", y="Portion of households with access", color="Limit type")
)
```

Remaining only 

For each DER type, we calculate the number of households with access based on grid limits, with "access" defined as all households having ability to host between 1.5 and 10 kW of PV.

```{r}
SCE_ICAaccess[["hhwacc_0_ICL_pct"]] <- ifelse(SCE_ICAaccess[["ICL_kWphh"]]>0, 1, 0)
PGE_ICAaccess[["hhwacc_0_ICL_pct"]] <- ifelse(PGE_ICAaccess[["ICL_kWphh"]]>0, 1, 0)

for (t in 1:length(at)) { 
  name <- paste0("hhwacc_",at[t],"_ICL_pct") 
  
  SCE_ICAaccess[[name]] <- ifelse(SCE_ICAaccess[["ICL_kWphh"]]>=as.numeric(at[t]), 1, SCE_ICAaccess[["ICL_kWphh"]]/as.numeric(at[t]))
  SCE_ICAaccess[[paste0("hhwacc_",at[t],"_ICL")]] <- SCE_ICAaccess[[name]]*SCE_ICAaccess$tothh_Cpoly
  
  PGE_ICAaccess[[name]] <- ifelse(PGE_ICAaccess[["ICL_kWphh"]]>=as.numeric(at[t]), 1, PGE_ICAaccess[["ICL_kWphh"]]/as.numeric(at[t]))
  PGE_ICAaccess[[paste0("hhwacc_",at[t],"_ICL")]] <- PGE_ICAaccess[[name]]*PGE_ICAaccess$tothh_Cpoly
}
```

Supplementary Figure S2
Visualizing results... just grid limits

```{r}
IOU_ICAaccess_grid <- 
  rbind(select(PGE_ICAaccess, IOU, tothh_Cpoly, GEOID, intersect(starts_with("hhwacc_"), ends_with("_pct"))),
        select(SCE_ICAaccess, IOU, tothh_Cpoly, GEOID, intersect(starts_with("hhwacc_"), ends_with("_pct")))) %>%
  melt(id=c("IOU", "tothh_Cpoly", "GEOID")) %>%
  mutate(hhwacc_Cpoly = value*tothh_Cpoly) %>%    # households per Cpoly, not percent
  group_by(IOU, GEOID, variable) %>%
  summarise(avgacc_bg_pct = mean(value, na.rm=TRUE), # avg percent access in bg
            tothh_bg = sum(tothh_Cpoly, na.rm=TRUE), # total number of households in block group
            hhwacc_bg_pct = sum(hhwacc_Cpoly, na.rm=TRUE)/sum(tothh_Cpoly, na.rm=TRUE))%>%
  mutate(variable = as.character(variable),
         threshold = sapply(strsplit(variable, "_"), function(x) x[2]), #change 3 to 2
         threshold = factor(ifelse(threshold=="0",">0",threshold), levels=c(">0",at)),
         # limit = sapply(strsplit(variable, "_"), function(x) x[2]),
         DER = factor(sapply(strsplit(variable, "_"), function(x) x[3]), levels=c("ICL")),  #change 4 to 3
         avgacc_bg_pct = as.numeric(avgacc_bg_pct))

# For the box grid plot
bg <- IOU_ICAaccess_grid %>% 
  filter(!is.na(hhwacc_bg_pct)) %>% 
  ggplot(aes(y=hhwacc_bg_pct)) + plotops_bggrid
bg

ggsave("figures/access/IOU_accessbg_ICL.png", width=8.5, height=6)
#ggsave("figures/access/IOU_accessbg_ICL.pdf", width=8.5, height=6)

IOU_accessbg_gr <- ggplot_build(bg)$data[[1]]

IOU_hhwacc_grid <- IOU_ICAaccess_grid %>%
  mutate(hhwacc_bg = hhwacc_bg_pct*tothh_bg) %>%
  mutate(DER = recode(DER, 'ICL'="Load")) %>%
  group_by(IOU, DER, threshold) %>%
  summarise(tothh = sum(tothh_bg, na.rm=TRUE),
            avgacc_pct = mean(avgacc_bg_pct, na.rm=TRUE),
            hhwacc_pct = sum(hhwacc_bg, na.rm=TRUE)/sum(tothh_bg, na.rm=TRUE)) %>%
  mutate(threshold = as.numeric(ifelse(as.character(threshold)==">0","0",as.character(threshold))))

# For the line and point grid plot
all <- IOU_hhwacc_grid %>% 
  ggplot(aes(x=threshold, y=hhwacc_pct)) + plotops_allgrid
all

# Combine the plots
plot_grid(all, bg, labels=c('a','b'), label_size=12, rel_widths=c(0.25,0.75), rel_heights=c(1,1))

ggsave("figures/access/IOU_access_ICL.png", width=11.5, height=6)
#ggsave("figures/access/IOU_access_ICL.pdf", width=11.5, height=6)
```

Merging back into main files

# ```{r}
# write.csv(SCE_ICAaccess, file = "circuitfiles/SCE_ICAaccess.csv", row.names=FALSE)
# write.csv(SCE_ICAaccess_grid, file = "circuitfiles/SCE_ICAaccess_bg_grid.csv", row.names=FALSE)
# write.csv(SCE_hhwacc_grid, file = "circuitfiles/SCE_ICAaccess_all_grid.csv", row.names=FALSE)
# ```
<!-- # -->
# ```{r}
# write.csv(PGE_ICAaccess, file = "circuitfiles/PGE_ICAaccess.csv", row.names=FALSE)
# write.csv(PGE_ICAaccess_grid, file = "circuitfiles/PGE_ICAaccess_bg_grid.csv", row.names=FALSE)
# write.csv(PGE_hhwacc_grid, file = "circuitfiles/PGE_ICAaccess_all_grid.csv", row.names=FALSE)
# ```


